---
title: "CSP571 Project Base Data Visualizations"
output: html_notebook
---

#Packages

```{r}
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)

if (!require("zoo")) install.packages("zoo")
library(zoo)

if (!require("dplyr")) install.packages("dplyr")
library(dplyr)

if (!require("tidyr")) install.packages("tidyr")
library(tidyr)
```

#Data Pull

```{r}
setwd("C:/Users/sydor/Downloads")

combined_data <- read.csv(file="FINAL Dataset - Combined_11_30 (3).csv", header=TRUE)
combined_data$Version <- as.Date(combined_data$Version, "%Y-%m-%d")
```

#Creates and returns plot for individual state

```{r}
graph_for_state <- function(state, policy, statistic, min_date, max_date) {
  state_data <- subset(combined_data, combined_data$Province_State == state)
  state_data <- subset(state_data, !is.na(state_data[,policy]))
  state_data <- subset(state_data, !is.na(state_data[,statistic]))
  
  #state_data[policy][state_data[policy]=="-"] <- NA
  #state_data[policy] <- na.locf(state_data[policy], na.rm = FALSE)
  
  lims <- as.Date(c(min_date, max_date), "%Y-%m-%d")
  
  state_plot <- ggplot(state_data, aes(x = Version, 
                                       y = state_data[,statistic], 
                                       col=as.factor(state_data[,policy]))) +

    #scale_y_continuous(limits = c(min(state_data$Mortality_Rate),max(state_data$Mortality_Rate))) +
    scale_x_date(date_breaks = "2 weeks", 
                 date_labels = "%b %d", 
                 limits = lims, 
                 #expand = c(0, 0)
                 ) +
    theme(axis.text.x = element_text(angle=45, 
                                     hjust = 1)) +
    geom_line() +
    geom_point(shape = 16) +
    labs(title=sprintf("%s Over Time in %s", statistic, state), 
         x = "Date", 
         y = statistic,
         col = policy) + 
    theme(legend.position="bottom", legend.title = element_text(face = "bold"))
  
  return(state_plot)
}
```

#Creates and returns model for individual state

```{r}
model_for_state <- function(state, policy, statistic, min_date, max_date) {
  
  state_data <- subset(combined_data, combined_data$Province_State == state)
  state_data <- subset(state_data, state_data$Version >= min_date)
  state_data <- subset(state_data, state_data$Version <= max_date)
  state_data <- subset(state_data, !is.na(state_data[,policy]))
  state_data <- subset(state_data, !is.na(state_data[,statistic]))
  
  loop <- unique(state_data[,policy])
  
  lm_list <- vector(mode = "list", length = length(loop))
  #test <- c()
  
  for(i in loop){
      policy_data <- subset(state_data, state_data[,policy]==i)
      
      lm <- lm(policy_data[,statistic] ~ policy_data$Version ,policy_data)
      #lm_list[[i+1]] <- summary(lm)
      lm_list[[i]] <- summary(lm)


  }
  return(lm_list)
}
```

#Creates and returns all models for every state, given a policy, statistic, and minimum number of observations

```{r}
model_all_state <- function(policy, statistic, min_num_obs) {
  cat("-----", policy, "-----")
  states <- unique(combined_data$Province_State)
  
  #Removing United States, Virgin Islands, and Washington D.C. from data
  states <- states[-c(44,47,50)]
  
  lm_list2 <- vector(mode = "list", length = length(states))
  
  for(i in 1:length(states)){
    cat("\n\n")
    cat(states[i], "(")
    cat(i)
    cat(")")
    state_data <- subset(combined_data, combined_data$Province_State == states[i])
    state_data <- subset(state_data, !is.na(state_data[,policy]))
    state_data <- subset(state_data, !is.na(state_data[,statistic]))
    
    severity <- unique(state_data[,policy])
    
    for(k in 1:length(severity)){
      cat("\n  Severity", severity[k])
      cat("\n    Num_obs:", nrow(subset(state_data, state_data[,policy] == severity[k])))
      if (is.na(subset(state_data, state_data[,policy] == severity[k])) | nrow(subset(state_data, state_data[,policy] == severity[k])) < min_num_obs){
        cat("\n      Removing severity", severity[k])
        cat(": only", nrow(subset(state_data, state_data[,policy] == severity[k])), "observation(s)")
        state_data <- subset(state_data, state_data[,policy] != severity[k])
      }
    }
    
    severity <- unique(state_data[,policy])
    cat("\n  Severities included:")
    
    if (length(severity)>0){
      lm_list <- vector(mode = "list", length = length(severity))
      for(j in 1:length(severity)){
        cat("", severity[j])
        
        deep <- vector(mode = "list", length = 2)
        policy_data <- subset(state_data, state_data[,policy] == severity[j])
        
        lm <- lm(policy_data[,statistic] ~ policy_data$Version, policy_data)
        deep[[1]] <- lm
        deep[[2]] <- severity[j]
        
        lm_list[[j]] <- deep
      }
    
    lm_list2[[i]] <- lm_list
    }
    
  }
  cat("\n\n------------------------------")
  return(lm_list2)
}
```

#Same as previous, but with optional date range

```{r}
model_all_state_with_date_lim <- function(policy, statistic, min_date, max_date, min_num_obs) {
  cat("-----", policy, "-----")
  
  states <- unique(combined_data$Province_State)
  states <- states[-c(44,47,50)]
  lm_list2 <- vector(mode = "list", length = length(states))
  
  for(i in 1:length(states)){
    cat("\n\n")
    cat(states[i], "(")
    cat(i)
    cat(")")
    state_data <- subset(combined_data, combined_data$Province_State == states[i])
    state_data <- subset(state_data, state_data$Version >= min_date)
    state_data <- subset(state_data, state_data$Version <= max_date)
    state_data <- subset(state_data, !is.na(state_data[,policy]))
    state_data <- subset(state_data, !is.na(state_data[,statistic]))
    
    severity <- unique(state_data[,policy])
    
    for(k in 1:length(severity)){
      cat("\n  Severity", severity[k])
      cat("\n    Num_obs:", nrow(subset(state_data, state_data[,policy] == severity[k])))
      
      if (is.na(subset(state_data, state_data[,policy] == severity[k])) | nrow(subset(state_data, state_data[,policy] == severity[k])) < min_num_obs){
        cat("\n      Removing severity", severity[k])
        cat(": only", nrow(subset(state_data, state_data[,policy] == severity[k])), "observation(s)")
        state_data <- subset(state_data, state_data[,policy] != severity[k])
      }
    }
    
    severity <- unique(state_data[,policy])
    cat("\n  Severities included:")
    
    if (length(severity)>0){
      lm_list <- vector(mode = "list", length = length(severity))
      
      for(j in 1:length(severity)){
        cat("", severity[j])
        
        deep <- vector(mode = "list", length = 2)
        policy_data <- subset(state_data, state_data[,policy] == severity[j])
        
        lm <- lm(policy_data[,statistic] ~ policy_data$Version, policy_data)
        deep[[1]] <- lm
        deep[[2]] <- severity[j]
        
        lm_list[[j]] <- deep
      }
    
    lm_list2[[i]] <- lm_list
    }
    
  }
  cat("\n\n------------------------------")
  return(lm_list2)
}
```

#Removes models with large p-values, given a list of models and a maximum allowed p-value

```{r}
remove_large_p <- function(state_models, max_p) {
  state_models_rm<-state_models
  for(i in 1:length(state_models)){
    cat("\n\nModel for state", i)
    
    for(j in 1:length(state_models[[i]])){
      if(length(state_models[[i]][[j]][[1]]) > 0){
        cat(":\n  Severity", state_models[[i]][[j]][[2]])
        cat("\n    P-value:", summary(state_models[[i]][[j]][[1]])$coefficients[2,4])
        
        if(is.na(summary(state_models[[i]][[j]][[1]])$coefficients[2,4])) {
          cat("\n    Removed: p-value is NA")
          state_models_rm[[i]][[j]][[1]] <- "Removed"
        } 
        else if(summary(state_models[[i]][[j]][[1]])$coefficients[2,4] > max_p){
          cat("\n    Removed: p-value is >", max_p)
          state_models_rm[[i]][[j]][[1]] <- "Removed"
        }
      }
    }
  }
  cat("\n\n------------------------------")
  return(state_models_rm)
}
```

#Gets coefficients from every model after removing unwanted models, given a list of wanted models

```{r}
get_coeff <- function(state_models_clean) {
  severity0<-c()
  severity1<-c()
  severity2<-c()
  severity3<-c()
  severity4<-c()
  
  for(i in 1:length(state_models_clean)){
    for(j in 1:length(state_models_clean[[i]])){
      if (length(state_models_clean[[i]][[j]][[1]]) > 0){
        if(state_models_clean[[i]][[j]][[1]] != "Removed"){
          if(state_models_clean[[i]][[j]][[2]] == 0){
            severity0<-c(severity0,summary(state_models_clean[[i]][[j]][[1]])$coefficients[2,1])
          }
          if(state_models_clean[[i]][[j]][[2]] == 1){
            severity1<-c(severity1,summary(state_models_clean[[i]][[j]][[1]])$coefficients[2,1])
          }
          if(state_models_clean[[i]][[j]][[2]] == 2){
            severity2<-c(severity2,summary(state_models_clean[[i]][[j]][[1]])$coefficients[2,1])
          }
          if(state_models_clean[[i]][[j]][[2]] == 3){
            severity3<-c(severity3,summary(state_models_clean[[i]][[j]][[1]])$coefficients[2,1])
          }
          if(state_models_clean[[i]][[j]][[2]] == 4){
            severity4<-c(severity4,summary(state_models_clean[[i]][[j]][[1]])$coefficients[2,1])
          }
        }
      }
    }
  }
  
  combined_all <- as.data.frame(cbind(s0=severity0,s1=severity1,s2=severity2,s3=severity3, s4=severity4))
  
  return(combined_all)
}
```

#Creates and returns boxplot of the spread of coefficients for each severity of a policy, given the list of coefficients and the statistic used (for the y-label)

```{r}
get_boxplot_coeff <- function(result, statistic) {
  result_long <- gather(result, Severity, y, factor_key=TRUE)

  ggplot(result_long, aes(x=Severity, y=y, fill=Severity)) + geom_boxplot() + labs(y=statistic)
}
```

#Calculates the mean of the coefficients for each severity of a policy, given the list of coefficients

```{r}
get_mean_coeff <- function(result){
  cat("\n\nMean of coefficients for each severity:\n")
  colMeans(result)
}
```

#Calls all functions needed for getting the coefficients

```{r}
do_all <- function(policy, statistic, max_p, min_num_obs){
  state_models <- model_all_state(policy, statistic, min_num_obs)
  state_models_clean <- remove_large_p(state_models, max_p)
  state_models_coeff <- get_coeff(state_models_clean)
  return (state_models_coeff)
}
```

#Same as previous, but with optional date range

```{r}
do_all_with_date_lim <- function(policy, statistic, min_date, max_date, max_p, min_num_obs){
  state_models <- model_all_state_with_date_lim(policy, statistic, min_date, max_date, min_num_obs)
  state_models_clean <- remove_large_p(state_models, max_p)
  state_models_coeff <- get_coeff(state_models_clean)
  return (state_models_coeff)
}
```

#Calling the functions

```{r, warning=FALSE}
policy <-"C2_Workplace.closing"
#C1_School.closing
#C2_Workplace.closing
#C3_Cancel.public.events
#C4_Restrictions.on.gatherings
#C5_Close.public.transport
#C6_Stay.at.home.requirements
#C7_Restrictions.on.internal.movement
#C8_International.travel.controls
#H3_Contact.tracing
#H6_Facial.Coverings

statistic <-"Daily_Confirmed"
#Daily_Confirmed
#Confirmed
#Daily_Deaths
#Deaths
#Mortality_Rate
#Hospitalization_Rate

#minimum number of observations required for model to be made
min_num_obs <- 10

#maximum p-value allowed for model - removed if higher
max_p <- 0.05

#getting models, plotting boxplot of coefficient spread, and getting mean of coefficients
return <- do_all(policy, statistic, max_p, min_num_obs)
get_boxplot_coeff(return, statistic)
get_mean_coeff(return)
```

#Calling the functions, including optional date range

```{r, warning=FALSE}
policy <-"H6_Facial.Coverings"
#C1_School.closing
#C2_Workplace.closing
#C3_Cancel.public.events
#C4_Restrictions.on.gatherings
#C5_Close.public.transport
#C6_Stay.at.home.requirements
#C7_Restrictions.on.internal.movement
#C8_International.travel.controls
#H3_Contact.tracing
#H6_Facial.Coverings

statistic <-"Daily_Confirmed"
#Daily_Confirmed
#Confirmed
#Daily_Deaths
#Deaths
#Mortality_Rate
#Hospitalization_Rate

#minimum number of observations required for model to be made
min_num_obs <- 10

#date limits optional
min_date <-"2020-04-12"
max_date <- "2020-11-30"

#maximum p-value allowed for model - removed if higher
max_p <- 0.05

#getting models, plotting boxplot of coefficient spread, and getting mean of coefficients
return <- do_all_with_date_lim(policy, statistic, min_date, max_date, max_p, min_num_obs)
get_boxplot_coeff(return, statistic)
get_mean_coeff(return)
```

#Plotting data with linear model line for testing purposes

```{r}
test <- model_for_state("Alabama", "C2_Workplace.closing", "Daily_Confirmed", "2020-04-12", "2020-11-30")
policy <- 3
summary(test[[policy]])

state_data <- subset(combined_data, combined_data$Province_State == "Alabama")
policy_data <- subset(state_data, state_data$C2_Workplace.closing==policy)

plot(policy_data$Version, policy_data$Daily_Confirmed, pch = 16, xlim= c(as.Date("2020-04-12"),as.Date("2020-11-30")), col = "red")
abline(test[[policy]], col = "blue")
```

#Getting model and plotting for particular state for testing purposes 

```{r}
model_for_state("Texas", "C1_School.closing", "Daily_Deaths", "2020-04-12", "2020-10-30")
graph_for_state("Florida", "C4_Restrictions.on.gatherings", "Daily_Confirmed", "2020-04-12", "2020-11-30")
```

```{r}
graph_for_state("Texas", "C1_School.closing", "Daily_Deaths", "2020-04-12", "2020-11-30")
```

