---
title: "CSP571 Project Base Data Visualizations"
output: html_notebook
---

#Packages

```{r}
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)

if (!require("zoo")) install.packages("zoo")
library(zoo)
```

#Data Pull

```{r}
setwd("D:/Kacper/School and Internship Work/Year 4 Sem 1/CSP 571 - Data Preperation and Analysis/PROJECT/COVID DATA/Visualizations")

combined_data <- read.csv(file="statistics_policies_combined_final.csv", header=TRUE)
combined_data$Version <- as.Date(combined_data$Version, "%Y-%m-%d")
```

#Function definition

```{r}
graph_for_state <- function(state, policy, statistic, min_date, max_date) {
  state_data <- subset(combined_data, combined_data$Province_State == state)
  
  state_data[policy][state_data[policy]=="-"] <- NA
  state_data[policy] <- na.locf(state_data[policy], na.rm = FALSE)
  state_data <- subset(state_data, state_data$Version >= min_date)
  state_data <- subset(state_data, state_data$Version <= max_date)
  
  lims <- as.Date(c(min_date, max_date), "%Y-%m-%d")
  
  state_plot <- ggplot(state_data, aes(x = Version, 
                                       y = state_data[,statistic], 
                                       col=as.factor(state_data[,policy]))) +

    #scale_y_continuous(limits = c(min(state_data$Mortality_Rate),max(state_data$Mortality_Rate))) +
    scale_x_date(date_breaks = "2 weeks", 
                 date_labels = "%b %d", 
                 limits = lims, 
                 #expand = c(0, 0)
                 ) +
    theme(axis.text.x = element_text(angle=45, 
                                     hjust = 1)) +
    geom_line() +
    geom_point(shape = 16) +
    labs(title=sprintf("%s Over Time in %s", statistic, state), 
         x = "Date", 
         y = statistic,
         col = policy) + 
    theme(legend.position="bottom", legend.title = element_text(face = "bold"))
  
  return(state_plot)
}
```

#Call the function using the following syntax: 
  graph_for_state(state, policy, statistic, min_date, max_date)
  
#where the input variables are defined as follows:

#state 
  Name of the state as a string.
  
#policy
  Name of the policy as a string. Possible values can be found in the source link.
  
#statistic
  Name of the statistic as a string. Possible values:
  - Daily_Confirmed
  - Confirmed
  - Daily_Deaths
  - Deaths
  - Recovered
  - Active
  - Incident_Rate
  - People_Tested
  - People_Hospitalized
  - Mortality_Rate
  - Testing_Rate
  - Hospitalization_Rate

#min_date:
  Minimum date for the x axis. The absolute minimum that the data allows for is "2020-04-12"

#max_date: 
  Maximum date for the x axis. The absolute maximum that the data allows for is "2020-08-31"

```{r}
graph_for_state("Texas", "C2_Workplace.closing", "Daily_Confirmed", "2020-04-12", "2020-11-30")
```

```{r}
graph_for_state("Arkansas", "C4_Restrictions.on.gatherings", "Daily_Confirmed", "2020-04-12", "2020-11-30")
```

```{r}
graph_for_state("Ohio", "C3_Cancel.public.events", "Mortality_Rate", "2020-04-12", "2020-08-31")
```